---
title: "Plotting with ggplot using some standard data sets"
output:
  html_document:
    df_print: paged
---

##Introduction

---

In this project I am plotting data from some standard data sets using standard ggplot functions. There is no specific order or grouping of these charts. Rather these are some of the plots that I used for my homework assignments at Carnegie Mellon University and 

```{r include = FALSE, message=FALSE}
#The R libraries listed below are used for this analysis. This block loads the libraries and reads data from the file
library(tidyverse)
library(ggplot2)
library(gridExtra)
#read data file from project data folder present in project working directory
telco_customer_churn_raw_data <- read_csv("data\\WA_Fn-UseC_-Telco-Customer-Churn.csv")

#Transform MultipleLines Variable
telco_customer_churn_transformed_data <- telco_customer_churn_raw_data %>%
     mutate(MultipleLines = ifelse(MultipleLines=="No phone service","No",MultipleLines))
#Transform OnlineSecurity Variable
telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(OnlineSecurity = ifelse(OnlineSecurity=="No internet service","No",OnlineSecurity))
#Transform OnlilneBackup Variable
telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(OnlineBackup = ifelse(OnlineBackup=="No internet service","No",OnlineBackup))
#Transform DeviceProtection Variable
telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(DeviceProtection = ifelse(DeviceProtection=="No internet service","No",DeviceProtection))
#Transform TechSupport Variable
telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(TechSupport = ifelse(TechSupport=="No internet service","No",TechSupport))
#Transform Streaming Movies Variable
telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(StreamingMovies = ifelse(StreamingMovies=="No internet service","No",StreamingMovies))
#Transform Streaming TV Variable
telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(StreamingTV = ifelse(StreamingTV=="No internet service","No",StreamingTV))
#Transform Senior Citizen Variable
telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(SeniorCitizen = ifelse(SeniorCitizen==1,"Yes","No"))
#Transform Churn Variable
telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(Churn = ifelse(Churn=="Yes",1,0))

telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(SeniorCitizen = as.factor(SeniorCitizen),
            gender = as.factor(gender),
            Partner = as.factor(Partner),
            Dependents = as.factor(Dependents),
            SeniorCitizen = as.factor(SeniorCitizen),
            PhoneService = as.factor(PhoneService),
            MultipleLines = as.factor(MultipleLines),
            InternetService = as.factor(InternetService),
            OnlineSecurity = as.factor(OnlineSecurity),
            OnlineBackup = as.factor(OnlineBackup),
            DeviceProtection = as.factor(DeviceProtection),
            TechSupport = as.factor(TechSupport),
            StreamingTV = as.factor(StreamingTV),
            StreamingMovies = as.factor(StreamingMovies),
            PaperlessBilling = as.factor(PaperlessBilling),
            Contract = as.factor(Contract),
            PaymentMethod = as.factor(PaymentMethod),
            Churn = as.factor(Churn))

telco_customer_churn_cleaned_data <- telco_customer_churn_transformed_data %>%
     mutate(TotalCharges = ifelse(is.na(TotalCharges),0,TotalCharges))
```

### 1. Count of categorical variable values

---

These plots were created using Customer Churn dataset. Here I have attempted to display the counts of all categorical variables in the data set using individual ggplot statements. Finally, I have knit all the plots into a grid using "grid.arrange".

```{r fig.width=15, fig.height=10}
#using ggplot here just to explore the functinalities in this package more
#Plot by SeniorCitizen
p1 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,PhoneService) %>%
     summarise(count = n()) %>%
     ggplot(aes(PhoneService,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("PhoneService distribution") +
    labs(
        x = "Phone Service",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by Gender
p2 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,MultipleLines) %>%
     summarise(count = n()) %>%
     ggplot(aes(MultipleLines,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("MultipleLines distribution") +
    labs(
        x = "Multiple Lines",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by Dependents
p3 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,InternetService) %>%
     summarise(count = n()) %>%
     ggplot(aes(InternetService,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("InternetService distribution") +
    labs(
        x = "Internet Service",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by OnlineSecurity
p4 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,OnlineSecurity) %>%
     summarise(count = n()) %>%
     ggplot(aes(OnlineSecurity,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("OnlineSecurity distribution") +
    labs(
        x = "Online Security",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by OnlineBackup
p5 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,OnlineBackup) %>%
     summarise(count = n()) %>%
     ggplot(aes(OnlineBackup,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("OnlineBackup distribution") +
    labs(
        x = "Online Backup",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by DeviceProtection
p6 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,DeviceProtection) %>%
     summarise(count = n()) %>%
     ggplot(aes(DeviceProtection,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("DeviceProtection distribution") +
    labs(
        x = "Device Protection",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by TechSupport
p7 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,TechSupport) %>%
     summarise(count = n()) %>%
     ggplot(aes(TechSupport,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("TechSupport distribution") +
    labs(
        x = "Tech Support",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by StreamingTV
p8 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,StreamingTV) %>%
     summarise(count = n()) %>%
     ggplot(aes(StreamingTV,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("StreamingTV distribution") +
    labs(
        x = "StreamingTV",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by StreamingMovies
p9 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,StreamingMovies) %>%
     summarise(count = n()) %>%
     ggplot(aes(StreamingMovies,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("StreamingMovies distribution") +
    labs(
        x = "Streaming Movies",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by PaperlessBilling
p10 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,PaperlessBilling) %>%
     summarise(count = n()) %>%
     ggplot(aes(PaperlessBilling,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("PaperlessBilling distribution") +
    labs(
        x = "Paperless Billing",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by Contract
p11 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,Contract) %>%
     summarise(count = n()) %>%
     ggplot(aes(Contract,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("Contract Type distribution") +
    labs(
        x = "Contract Type",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by PaymentMethod
p12 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,PaymentMethod) %>%
     summarise(count = n()) %>%
     ggplot(aes(PaymentMethod,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("PaymentMethod distribution") +
    labs(
        x = "Payment Method",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )

grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,p12)
```

### 2. ggplots with some animation

---

In this section I attempted to add some animation to ggplots using gganimate package. Such plots may not be very useful for Exploratory Data Analyses, but they could be a cool option to include on a website to show trends.

```{r echo=FALSE, message=FALSE, height=6, width = 6}

library(ggExtra)
library(gganimate)
library(gapminder)
library(gifski)
library(magick)

land_data <- read.csv("data\\landdata-states.csv")
land_data_qtr1 <- land_data %>%
                    filter(Qrtr ==1)

ggplot(land_data_qtr1,aes(Land.Share..Pct.,Land.Value)) +
     geom_point(aes(color = region)) +
     theme_bw()+
     labs(
          title = 'Land value vs Land share % in House value: {closest_state}',
          x = "Land Share Percentage",
          y = "Land Value"
     ) + 
     theme(
        plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
        axis.title.x = element_text(face = "bold", size = 14),
        axis.title.y = element_text(face = "bold", size = 14),
        axis.text.x = element_text(face = "bold",size = 12),
        axis.text.y = element_text(face = "bold", size = 12),
        legend.position = "none"
     ) +
     transition_states(Year, state_length = 30) +
     ease_aes('sine-in-out')
```

---

<br>
```{r echo=FALSE, message=FALSE, height = 6, width=6}
land_data_qtr1 %>%
     filter(region=="West") %>%
     ggplot(aes(State, Land.Share..Pct.)) +
          geom_col(aes(fill="red")) +
          theme_bw() +
          labs(
               title = 'Land share % in total house value in Western US: {closest_state}',
               x = "State",
               y = "Landshare percentage in house value"
          ) + 
          theme(
             plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
             axis.title.x = element_text(face = "bold", size = 14),
             axis.title.y = element_text(face = "bold", size = 14),
             axis.text.x = element_text(face = "bold",size = 12),
             axis.text.y = element_text(face = "bold", size = 12),
             legend.position = "none"
          ) +
          transition_states(Year, state_length = 30)

```
