---
title: "Plotting with ggplot using some standard data sets"
output: html_notebook
---

```{r include = FALSE, message=FALSE}
#The R libraries listed below are used for this analysis. This block loads the libraries and reads data from the file
library(tidyverse)
library(ggplot2)
#read data file from project data folder present in project working directory
telco_customer_churn_raw_data <- read_csv("data\\WA_Fn-UseC_-Telco-Customer-Churn.csv")

#Transform MultipleLines Variable
telco_customer_churn_transformed_data <- telco_customer_churn_raw_data %>%
     mutate(MultipleLines = ifelse(MultipleLines=="No phone service","No",MultipleLines))
#Transform OnlineSecurity Variable
telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(OnlineSecurity = ifelse(OnlineSecurity=="No internet service","No",OnlineSecurity))
#Transform OnlilneBackup Variable
telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(OnlineBackup = ifelse(OnlineBackup=="No internet service","No",OnlineBackup))
#Transform DeviceProtection Variable
telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(DeviceProtection = ifelse(DeviceProtection=="No internet service","No",DeviceProtection))
#Transform TechSupport Variable
telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(TechSupport = ifelse(TechSupport=="No internet service","No",TechSupport))
#Transform Streaming Movies Variable
telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(StreamingMovies = ifelse(StreamingMovies=="No internet service","No",StreamingMovies))
#Transform Streaming TV Variable
telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(StreamingTV = ifelse(StreamingTV=="No internet service","No",StreamingTV))
#Transform Senior Citizen Variable
telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(SeniorCitizen = ifelse(SeniorCitizen==1,"Yes","No"))
#Transform Churn Variable
telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(Churn = ifelse(Churn=="Yes",1,0))

telco_customer_churn_transformed_data <- telco_customer_churn_transformed_data %>%
     mutate(SeniorCitizen = as.factor(SeniorCitizen),
            gender = as.factor(gender),
            Partner = as.factor(Partner),
            Dependents = as.factor(Dependents),
            SeniorCitizen = as.factor(SeniorCitizen),
            PhoneService = as.factor(PhoneService),
            MultipleLines = as.factor(MultipleLines),
            InternetService = as.factor(InternetService),
            OnlineSecurity = as.factor(OnlineSecurity),
            OnlineBackup = as.factor(OnlineBackup),
            DeviceProtection = as.factor(DeviceProtection),
            TechSupport = as.factor(TechSupport),
            StreamingTV = as.factor(StreamingTV),
            StreamingMovies = as.factor(StreamingMovies),
            PaperlessBilling = as.factor(PaperlessBilling),
            Contract = as.factor(Contract),
            PaymentMethod = as.factor(PaymentMethod),
            Churn = as.factor(Churn))
```

### 1. Count of categorical variable values

---

These plots were created using Customer Churn dataset. Here I have attempted to display the counts of all categorical variables in the data set using individual ggplot statements. Finally, I have knit all the plots into a grid using "grid.arrange".

There is also a shorter way of doing this, which I have attempted in the next section.

```{r fig.width=15, fig.height=10}
#using ggplot here just to explore the functinalities in this package more
#Plot by SeniorCitizen
p1 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,PhoneService) %>%
     summarise(count = n()) %>%
     ggplot(aes(PhoneService,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("PhoneService distribution") +
    labs(
        x = "Phone Service",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by Gender
p2 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,MultipleLines) %>%
     summarise(count = n()) %>%
     ggplot(aes(MultipleLines,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("MultipleLines distribution") +
    labs(
        x = "Multiple Lines",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by Dependents
p3 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,InternetService) %>%
     summarise(count = n()) %>%
     ggplot(aes(InternetService,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("InternetService distribution") +
    labs(
        x = "Internet Service",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by OnlineSecurity
p4 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,OnlineSecurity) %>%
     summarise(count = n()) %>%
     ggplot(aes(OnlineSecurity,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("OnlineSecurity distribution") +
    labs(
        x = "Online Security",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by OnlineBackup
p5 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,OnlineBackup) %>%
     summarise(count = n()) %>%
     ggplot(aes(OnlineBackup,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("OnlineBackup distribution") +
    labs(
        x = "Online Backup",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by DeviceProtection
p6 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,DeviceProtection) %>%
     summarise(count = n()) %>%
     ggplot(aes(DeviceProtection,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("DeviceProtection distribution") +
    labs(
        x = "Device Protection",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by TechSupport
p7 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,TechSupport) %>%
     summarise(count = n()) %>%
     ggplot(aes(TechSupport,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("TechSupport distribution") +
    labs(
        x = "Tech Support",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by StreamingTV
p8 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,StreamingTV) %>%
     summarise(count = n()) %>%
     ggplot(aes(StreamingTV,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("StreamingTV distribution") +
    labs(
        x = "StreamingTV",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by StreamingMovies
p9 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,StreamingMovies) %>%
     summarise(count = n()) %>%
     ggplot(aes(StreamingMovies,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("StreamingMovies distribution") +
    labs(
        x = "Streaming Movies",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by PaperlessBilling
p10 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,PaperlessBilling) %>%
     summarise(count = n()) %>%
     ggplot(aes(PaperlessBilling,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("PaperlessBilling distribution") +
    labs(
        x = "Paperless Billing",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by Contract
p11 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,Contract) %>%
     summarise(count = n()) %>%
     ggplot(aes(Contract,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("Contract Type distribution") +
    labs(
        x = "Contract Type",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )
#Plot by PaymentMethod
p12 <- telco_customer_churn_cleaned_data %>%
     group_by(Churn,PaymentMethod) %>%
     summarise(count = n()) %>%
     ggplot(aes(PaymentMethod,count,fill = Churn))+
     geom_bar(stat = "identity", position = "dodge")+
     coord_flip()+
     ggtitle("PaymentMethod distribution") +
    labs(
        x = "Payment Method",
        y = "Frequency"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.background = element_rect(
            fill = "white",
            colour = "lightgrey",
            size = 0.75),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_line(colour = "#C4DFE6", size = 0.05)
    )

grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,p12)
```
